Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-12-31 article.tex
All sections finished; ready for co-author review after final checks

\abstract
{
In this study, we explore the efficacy of Jacobian-free Newton-Krylov methods within the context of finite-volume solid mechanics.
Traditional Newton-based approaches to solving nonlinear systems typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
The Jacobian-free Newton-Krylov method circumvents this by employing Krylov subspace iterative solvers, such as GMRES, in conjunction with a Newton iteration scheme that approximates the action of the Jacobian through finite difference evaluations.
A further potential advantage of the Jacobian-free Newton-Krylov method is that it is readily applicable to existing segregated finite volume frameworks, where forming and storing the exact Jacobian would require major code refactoring.
%This approach promises significant computational savings, especially for large-scale, complex simulations prevalent in solid mechanics.
This article research systematically evaluates the performance of Jacobian-free Newton-Krylov methods by benchmarking them against conventional segregated methods on a suite of benchmark cases of varying geometric dimension, geometric nonlinearity, dynamic response, and material behaviour.
Key metrics such as computational cost, memory and robustness are analysed.
Additionally, we investigate the impact of various solution algorithm choices, such as preconditioning strategy, on the efficiency of the Jacobian-free Newton-Krylov method.
Our findings indicate that Jacobian-free Newton-Krylov methods can achieve \hl{comparable/superior/XXX convergence behaviour} relative to traditional segregated methods, particularly in cases where YYYY.
\hl{Summarise key findings: time, memory, important choices, JFNK vs SEG vs FE, ...}
The results suggest that Jacobian-free Newton-Krylov methods are promising for advancing finite-volume solid mechanics simulations and are particularly attractive for existing segregated frameworks where minimal code changes would be required to exploit openly available Jacobian-free Newton-Krylov implementations.
The described implementations are made publicly available in the solids4foam toolbox for OpenFOAM, allowing the community to examine, extend and compare the procedures with the our codes.
%offering a viable pathway for enhancing computational efficiency and scalability.
%EMPHASISE: easy to extend segregated frameworks, on contrast to exact Jacobian methoods.
%This study provides critical insights and practical guidelines for implementing JFNK methods in engineering and scientific applications.
%Key FV points:
%- many FV codes were developed around a segregated solution procedure, which requires significant effort to extend to a full Newton method, e.g. in terms of Jacobian assembly, storage, and linear system solution.
%- this paper examines Jacobian-free Newton-Krylov as a straight-forward extension of the segregated approach, without the need for a full Jacobian based method.
%- compact approximate Jacobian for preconditioner (more compact than FE approach)
%- implemented in OpenFOAM, and code and cases are made publicly available.
}

\keywords{Jacobian-free Newton-Krylov, Finite volume method, GMRES, OpenFOAM}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Paper outline:
%
%Intro
%- FV is of interest for CSM
%- Most methods use segregated algorithms, stemming from CFD algorithms
%- Extension to block-coupled Newton methods is not easy, in terms of derivation, and code refactoring (matrix storage, extended stencil, linear solver, etc.).
%- JFNK promises the performance of Newton methods, but without the need to form the full Jacobian. Define JFNK method. Hence, an existing segregated code can easily be adapted to use JFNK without major refactoring, albeit this process is made easier by the availability of open-source JFNK implementations.
%- this paper examines JFNK for linear and nonlinear FV CSM procedures, where the compact stencil approximate Jacobian is used for preconditioning.
Finite volume formulations for solid mechanics are heavily influenced by their fluid mechanics counterparts, favouring fully explicit \citep{Trangenstein1991, Kluth2010, Lee2013, Haider2017} or segregated implicit \citep{Demirdzic1988, Fryer1991, Demirdzic1995, Jasak2000, Tukovic2013, Cardiff2017, Tukovic2018, Batistic2022} methods.
Segregated approaches, where the governing momentum equation is temporarily decomposed into scalar component equations, offer memory efficiency and simplicity of implementation, but the outer coupling Picard iterations often suffer from slow convergence.
Explicit formulations are straightforward to implement and offer superior robustness but are only efficient for high-speed dynamics, where the physics requires small time increments.
In contrast, the finite element community commonly employs Newton-Raphson-type solution algorithms, which necessitate repeated assembly of the Jacobian matrix and solution of the resulting block-coupled non-diagonally dominant linear system.
A disadvantage of traditional Newton-based approaches is that they typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
A further disadvantage from a finite volume perspective is that extending existing code frameworks from segregated algorithms to coupled Newton-Raphson-type approaches is challenging in terms of the required assembly, storage, and solution of the resulting block-coupled system.
In addition, the derivation of the true Jacobian matrix is non-trivial.
Consequently, similar block-coupled solution finite volume methods are rare in the literature \citep{Das2011, Cardiff2016, Castrillo2024}.
The motivation of the current work is to seek (or exceed) the robustness and efficiency of block-coupled Newton-Raphons approaches in a way that can be easily incorporated into existing segregated solution frameworks.
To this end, the current article examines the efficacy of \emph{Jacobian-free} Newton-Krylov methods, where the quadratic convergence of Newton methods can potentially be achieved without deriving, assembling and storing the exact Jacobian.

Jacobian-free Newton-Krylov methods circumvent the need for the Jacobian matrix by combining the Newton-Raphson method with Krylov subspace iterative linear solvers, such as GMRES, and noticing that such Krylov solvers do not explicitly require the Jacobian matrix.
Instead, only the action of the Jacobian matrix on a solution-type vector is required.
The key step in Jacobian-free Newton-Krylov methods is the approximation of products between the Jacobian matrix and a vector using the finite difference method; that is
\begin{eqnarray}
	\bb{J} \bb{v} \approx \frac{\bb{F}(\bb{x} + \epsilon \bb{v}) - \bb{F}(\bb{x})}{\epsilon}
\end{eqnarray}
where $\mathbf{J}$ is the Jacobian matrix, $\mathbf{x}$ is the current solution vector (e.g. nodal displacements), $\mathbf{v}$ is a vector (e.g., from a Krylov subspace), and $\epsilon$ is a small scalar perturbation.
%Determining the appropriate value for $\epsilon$ requires balancing the truncation error of the finite difference approximation and round-off (numerical precision) error.
With an appropriate choice of $\epsilon$ (balancing truncation and round-off errors), the characteristic quadratic convergence of Newton methods can be achieved without the Jacobian, hence the modifier \emph{Jacobian-free}.
This approach promises significant memory savings over Jacobian-based methods, especially for large-scale problems, but also potentially for execution time, with appropriate choice of solution components.

@@

They found that the lower-order compact stencil preconditioner gave the best performance when combined with the GMRES linear solver and ILU(k) preconditioner.

@@

They also found a multigrid approach to be more efficient than ILU for preconditioning, where the implementation was based on the PETSc \citep{PETSc} package.

@@

A particularly attractive application of the low-order-discretisation preconditioned Newton-Krylov is for higher-order discretisations; that is, spatial discretisation of order greater than two.

@@

GMRES was the adopted linear solver with the ILU(k) preconditioner and fill-in values (k) ranging from 2 to 4. 

@@

Although they are increasingly seen in finite volume computational fluid dynamics, Jacobian-free Newton-Krylov methods have yet to be used for finite volume solid mechanics.
This article aims to address this point by assessing the efficacy of Jacobian-free Newton-Krylov cell-centred finite volume formulations applied to static, dynamic, linear and nonlinear solid mechanics problems.
Particular focus is given to the use of a compact-stencil approximate Jacobian preconditioner, inspired from existing segregated solid mechanics procedures, and ease of implementation into an existing segregated finite volume framework -- OpenFOAM \citep{Weller1998}.
%Novel in this paper:
%- JKNK for FVM CSM/CSD
%- focus on easy-of-implementation in an existing segregated code framework.
%- explore compact approximate Jacobian as a suitable preconditioner for linear and nonlinear solid mechanics;
%- explore multigrid as a preconditioner for JFNK
The remainder of the paper is structured as follows:
Section 2 summarises a typical solid mechanics mathematical model and its cell-centred finite volume discretisation.
Section 3 presents the solution algorithms, starting with the classic segregated solution algorithm, followed by the proposed Jacobian-free Newton-Krylov solution algorithm.
The performance of the proposed Jacobian-free Newton-Krylov approach is compared with the segregated approach on several varying benchmark cases in Section 4, where the effect of several factors are examined, including problem dimension, mesh, material model, nonlinear geometry, choice of preconditioner, and other solution parameter.
Finally, the article ends with a summary of the main conclusions of the work.

@@

The governing equations are complemented by boundary conditions, with three types considered here: prescribed displacement, prescribed traction, and symmetry.
The definition of the engineering stress ($\bb{\sigma}_s$) and true stress ($\bb{\sigma}$) in Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL} is given by a chosen mechanical law.
Three mechanical laws are considered in this work, as briefly outlined in Appendix \ref{app:mechLaws}: linear elasticity (Hooke's law), neo-Hookean hyperelasticity, and neo-Hookean $J_2$ hyperelastoplasticity.

@@

\paragraph{Case 6: Elastic plate behind a rigid cylinder}
The final test case, introduced by \citet{Turek2006}, extends the classic flow around a cylinder in a channel problem \citep{Ferziger2002} into a well-established fluid-solid interaction benchmark.
The \texttt{FSI3} variant of the case examined here (Figure \ref{fig:hronTurek-mesh}) consists of a horizontal channel (0.41 m in height, 2.5 m in length) with a rigid cylinder of radius 0.05 m, where the cylinder centre is 0.2 m from the bottom and inlet (left) boundaries.
A parabolic velocity is prescribed at the inlet velocity with a mean value of 2 m/s.
A St.~Venant-Kirchhoff hyperelastic plate ($E = 5.6$ MPa, $\nu = 0.4$) of 0.35 m in length and 0.02 m in height is attached to the right-hand side of the rigid cylinder.
The fluid model is assumed to be isothermal, incompressible, and laminar and adopts the segregated PIMPLE solution algorithm.
The fluid's kinematic viscosity is 0.001 m$^2$/s and density is 1000 kg/m$^3$, while the solid's density is $10\,000$ kg/m$^3$.
The interface quasi-Newton coupling approach with inverse Jacobian from a least-squares model \cite{Degroote2009} is employed for the fluid-solid interaction coupling.
The fluid-solid interface residual is reduced by four orders of magnitude within each time step.
Further details of the fluid-solid interaction procedure are found in \citet{Tukovic2018a}.
%\begin{figure}[htbp]
%   \centering
%%	\subfigure[Geometry (taken from \citet{Tukovic2018a})]
%%	{
%	   \includegraphics[width=0.7\textwidth]{figures/hronTurek-geometry} 
%%   	}
%   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
%   \label{fig:hronTurek-geometry}
%\end{figure}
Three successively refined quadrilateral meshes are employed:
The fluid region meshes have $1\,252$, $5\,008$, and $20\,032$ cells, while the solid region meshes have 156,  624 and $2\,496$ cells.
The total simulation time is 20 s, and the time-step size is 0.5 ms. 
\begin{figure}[htbp]
   \centering
	\subfigure[Mesh containing $5\,336$ cells in the fluid region and 630 cells in the solid region, where a closeup of the plate is shown in the right]
	{
   		\includegraphics[width=0.7\textwidth]{figures/hronTurek-mesh}
	}
	\subfigure[Close-up of the mesh around the plate]
	{
   		\includegraphics[width=0.4\textwidth]{figures/hronTurek-mesh-zoom}  
   	}
   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
   \label{fig:hronTurek-mesh}
\end{figure}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Resource Requirements and Robustness}
\label{sec:resource_requirements}
%%--------------------------------------------------------------------------------------------------------------------%%
This section compares resource requirements and the robustness of the Jacobian-free Newton-Krylov and segregated approaches on the cases presented in Section \ref{sec:accuracy}.
Specifically, time and memory requirements are analysed, and whether the solution approach converged for all meshes and loading steps.
%In some cases, time and memory requirements from finite element software \hl{Abaqus} are given for reference.
In all cases, the Jacobian-free Newton-Krylov approach used the generalised minimal residual method (GMRES) linear solver with the direct LU preconditioner for 2-D and the Hypre \emph{BoomerAMG} multigrid preconditioner for 3-D, unless stated otherwise.
A restart value of 30 is used for the GMRES solver, where the \emph{loose} variant of the GMRES solver is used, which retains a small number of extra solution vectors across restarts (2 in this case).
In contrast, The segregated approach used the conjugate gradient linear solver and the Hypre \emph{BoomerAMG} multigrid preconditioner for all cases.
Time and memory requirements are implementation and hardware-specific; nonetheless, it is insightful to see the relative performances of the Jacobian-free Newton-Krylov and segregated approaches on the same hardware and within the same implementation framework.
Clock times and memory usage (measured with the GNU time utility) were generated using a Mac Studio with an M2 Ultra CPU on one CPU core, where the code was built with the Clang compiler (version 16.0.0).
%\hl{Present data for one hardware first, and give comparison after.}
Examination of multi-CPU-core parallelisation is left to Section \ref{sec:parallelisation}.
% where several mesh densities are examined.
%Time and memory requirements from the classic segregated approach are given for comparison, along with the requirements from commercial finite element software Abaqus (version XXX): \hl{ask Dylan to create the Abaqus results when we are happy the paper will not change}.
%The finite element approach uses a coupled solution algorithm and direct linear solver, where a Newton-Raphson method is used for resolving nonlinearities.
%\hl{Take care when comparing times between different hardware}

@@

Regarding memory usage, below approximately $50\,000$ degrees of freedom, the memory usage is the same for all three approaches (approximately $80$ MB), corresponding to the solver's minimum memory overhead.
For greater numbers of degrees of freedom, the LU approach uses the greatest amount of memory for both 3-D cases (idealised ventricle, elliptic plate), while the ILU(5) uses the least.
In contrast, the LU approach uses the smallest amount of memory for the 2-D case (membrane i) mesh densities examined.
The rate of memory increase is seen to be much steeper for the LU approach on the 3-D cases than the multigrid and ILU(5) approaches, with the LU results for the finest meshes not shown as they required more than the maximum available memory (64 GB).
%\begin{itemize}
%	\item \hl{Ventricle finesh mesh: ILU5 diverges and LU requires too much memory}
%	\item \hl{Elliptic plate finest mesh: LU requires too much memory}
%	\item \hl{Comment on the results: e.g. fastest, slowest, most/least iterations, etc}
%	\item \hl{Provide insights, if any: e.g. why is one faster or slower}
%	\item \hl{Conclusion: which preconditioner if 'best' for which situation: probably LU is best for 'small' cases and multigrid for 'large' cases}
%\end{itemize}
\begin{figure}[htbp]
   \centering
	\subfigure[Clock time]
	{
   		\includegraphics[width=0.48\textwidth]{figures/time}
	}
	\subfigure[Maximum memory]
	{
   		\includegraphics[width=0.48\textwidth]{figures/memory}  
   	}
   \caption{The clock times and maximum memory usage for three preconditioning approaches (multigrid, ILU(5), and LU) as a function of degrees of freedom}
   \label{fig:times_memory}
\end{figure}
%\begin{table}[htb]
%	\centering
%		\begin{tabular}{lllll}
%			\hline
%			Case & Number of Cells & Preconditioner & Time (in s) & Memory (in MB)  \\
%			\hline 
%			Case X & 1,234 & LU & 345 & 1,234  \\
%			Case X & 12,345 & LU & 345 & 1,234  \\
%			Case X & 123,456 & LU & 345 & 1,234  \\
%			Case X & 1,234 & ILU(0) & 345 & 1,234  \\
%			Case X & 12,345 & ILU(0) & 345 & 1,234  \\
%			Case X & 123,456 & ILU(0) & 345 & 1,234  \\
%			Case X & 1,234 & Multigrid & 345 & 1,234  \\
%			Case X & 12,345 & Multigrid & 345 & 1,234  \\
%			Case X & 123,456 & Multigrid & 345 & 1,234  \\
%			Case Y & 1,234 & LU & 345 & 1,234  \\
%			Case Z & 1,234 & LU & 345 & 1,234  \\
%			\hline
%		\end{tabular}
%	\caption{Execution times and maximum memory usage for different preconditioners on varying cases and mesh densities \hl{Try other value of N for ILU?}}
%	\label{tab:preconditioner}
%\end{table}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Parallelisation}
\label{sec:parallelisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this final analysis, the multi-CPU-core parallel scaling performance of the proposed Jacobian-free Newton Krylov method is compared with the segregated approach.
A \emph{strong} scaling study is performed, where the clock time to solve a fixed-size problem is measured as the number of CPU cores is increased.
The elliptic plate (3-D, linear elastic) mesh with $2\,438\,242$ cells is chosen for the scaling analysis.
Parallelisation adopts the domain decomposition approach, where the mesh is decompose into one sub-domain for each CPU core.
In the current work, the Scotch decomposition approach \citep{Pellegrini2012} is employed.
The Jacobian-free Newton Krylov approach uses the GMRES linear solver with the multigrid preconditioner, while the segregated approach uses the conjugate gradient solver with the incomplete (zero in fill) Cholesky preconditioner.
The cases are run on the Meluxina high-performance computing system, where each standard computing node contains $2\times$ AMD EPYC Rome 7H12 64c 2.6GHz CPUs with 512 GB of memory.
In the current study, the number of CPU cores are varied from 1 to $1\,048$ ($\left[ 1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1\,048 \right]$).
In the ideal case, the speedup should double when the number of cores is doubled; in reality, inter-CPU-core communication reduces the parallel scaling efficiency below the ideal.

>>>>

\abstract
{
This study investigates the efficacy of Jacobian-free Newton-Krylov methods in finite-volume solid mechanics. Traditional Newton-based approaches require explicit Jacobian matrix formation and storage, which can be computationally expensive and memory-intensive.
In contrast, Jacobian-free Newton-Krylov methods approximate the Jacobian’s action using finite differences, combined with Krylov subspace solvers such as GMRES, enabling seamless integration into existing segregated finite-volume frameworks without major code refactoring.
This work benchmarks the performance of a compact-stencil Jacobian-free Newton-Krylov method against a conventional segregated approach on a suite of test cases, encompassing varying geometric dimensions, nonlinearities, dynamic responses, and material behaviours.
Key metrics, including computational cost, memory efficiency, and robustness, are evaluated, along with the influence of preconditioning strategies and stabilisation scaling.
Results show that the proposed Jacobian-free Newton-Krylov method outperforms the segregated approach in all linear and nonlinear elastic cases, achieving order-of-magnitude speedups in many instances; however, divergence is observed in elastoplastic cases, highlighting areas for further development.
It is found that preconditioning choice significantly impacts performance: a LU direct solver is fastest in small to moderately-sized cases, while a multigrid method is more effective for larger problems.
The findings demonstrate that Jacobian-free Newton-Krylov methods are promising for advancing finite-volume solid mechanics simulations, particularly for existing segregated frameworks where minimal modifications enable their adoption.
The described implementations are available in the solids4foam toolbox for OpenFOAM, inviting the community to explore, extend, and compare these procedures.
}

\keywords{Jacobian-free Newton-Krylov, Finite volume method, GMRES, solids4foam, OpenFOAM}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Paper outline:
%
%Intro
%- FV is of interest for CSM
%- Most methods use segregated algorithms, stemming from CFD algorithms
%- Extension to block-coupled Newton methods is not easy, in terms of derivation, and code refactoring (matrix storage, extended stencil, linear solver, etc.).
%- JFNK promises the performance of Newton methods, but without the need to form the full Jacobian. Define JFNK method. Hence, an existing segregated code can easily be adapted to use JFNK without major refactoring, albeit this process is made easier by the availability of open-source JFNK implementations.
%- this paper examines JFNK for linear and nonlinear FV CSM procedures, where the compact stencil approximate Jacobian is used for preconditioning.
Finite volume formulations for solid mechanics are heavily influenced by their fluid mechanics counterparts, favouring fully explicit \citep{Trangenstein1991, Kluth2010, Lee2013, Haider2017} or segregated implicit \citep{Demirdzic1988, Fryer1991, Demirdzic1995, Jasak2000, Tukovic2013, Cardiff2017, Tukovic2018, Batistic2022} methods.
Segregated approaches, where the governing equations are temporarily decomposed into scalar component equations, offer memory efficiency and simplicity of implementation, but the outer coupling Picard iterations often suffer from slow convergence.
Explicit formulations are straightforward to implement and offer superior robustness but are only efficient for high-speed dynamics, where the physics requires small time increments.
In contrast, the finite element community commonly employs Newton-Raphson-type solution algorithms, which necessitate repeated assembly of the Jacobian matrix and solution of the resulting block-coupled non-diagonally dominant linear system.
A disadvantage of traditional Newton-based approaches is that they typically require explicit formation and storage of the Jacobian matrix, which can be computationally expensive and memory-intensive.
A further disadvantage from a finite volume perspective is that extending existing code frameworks from segregated algorithms to coupled Newton-Raphson-type approaches is challenging in terms of the required assembly, storage, and solution of the resulting block-coupled system.
In addition, the derivation of the true Jacobian matrix is non-trivial.
Consequently, similar block-coupled solution finite volume methods are rare in the literature \citep{Das2011, Cardiff2016, Castrillo2024}.
The motivation of the current work is to seek the robustness and efficiency of block-coupled Newton-Raphons approaches in a way that can be easily incorporated into existing segregated solution frameworks.
To this end, the current article examines the efficacy of \emph{Jacobian-free} Newton-Krylov methods, where the quadratic convergence of Newton methods can potentially be achieved without deriving, assembling and storing the exact Jacobian.

Jacobian-free Newton-Krylov methods circumvent the need for the Jacobian matrix by combining the Newton-Raphson method with Krylov subspace iterative linear solvers, such as GMRES, and noticing that such Krylov solvers do not explicitly require the Jacobian matrix.
Instead, only the action of the Jacobian matrix on a solution-type vector is required.
The key step in Jacobian-free Newton-Krylov methods is the approximation of products between the Jacobian matrix and a vector using the finite difference method; that is
\begin{eqnarray}
	\bb{J} \bb{v} \approx \frac{\bb{F}(\bb{x} + \epsilon \bb{v}) - \bb{F}(\bb{x})}{\epsilon}
\end{eqnarray}
where $\bb{J}$ is the Jacobian matrix, $\mathbf{x}$ is the current solution vector (e.g. nodal displacements), $\mathbf{v}$ is a vector (e.g., from a Krylov subspace), and $\epsilon$ is a small scalar perturbation.
%Determining the appropriate value for $\epsilon$ requires balancing the truncation error of the finite difference approximation and round-off (numerical precision) error.
With an appropriate choice of $\epsilon$ (balancing truncation and round-off errors), the characteristic quadratic convergence of Newton methods can be achieved without the Jacobian, hence the modifier \emph{Jacobian-free}.
This approach promises significant memory savings over Jacobian-based methods, especially for large-scale problems, but also potentially for execution time, with appropriate choice of solution components.

@@

They found that the lower-order compact stencil preconditioner gave the best performance when combined with the GMRES linear solver and ILU(N) preconditioner.

@@

They also found a multigrid approach to be more efficient than ILU(N) for preconditioning, where the implementation was based on the PETSc \citep{PETSc} package.

@@

A particularly attractive application of the low-order preconditioning approach is for higher-order discretisations; that is, spatial discretisation of order greater than two.

@@

GMRES was the adopted linear solver with the ILU(N) preconditioner and fill-in values (N) ranging from 2 to 4. 

@@

Although they are increasingly seen in finite volume computational fluid dynamics, Jacobian-free Newton-Krylov methods have yet to be used for finite volume solid mechanics.
This article aims to address this point by assessing the efficacy of Jacobian-free Newton-Krylov cell-centred finite volume formulations applied to static, dynamic, linear and nonlinear solid mechanics problems.
Particular focus is given to the use of a compact-stencil approximate Jacobian preconditioner, inspired from existing segregated solid mechanics procedures, and ease of implementation into an existing segregated finite volume framework -- OpenFOAM \citep{Weller1998}.
%Novel in this paper:
%- JKNK for FVM CSM/CSD
%- focus on easy-of-implementation in an existing segregated code framework.
%- explore compact approximate Jacobian as a suitable preconditioner for linear and nonlinear solid mechanics;
%- explore multigrid as a preconditioner for JFNK
The remainder of the paper is structured as follows:
Section 2 summarises a typical solid mechanics mathematical model and its cell-centred finite volume discretisation.
Section 3 presents the solution algorithms, starting with the classic segregated solution algorithm, followed by the proposed Jacobian-free Newton-Krylov solution algorithm.
The performance of the proposed Jacobian-free Newton-Krylov approach is compared with the segregated approach on several varying benchmark cases in Section 4, where the effect of several factors are examined. %, including problem dimension, mesh, material model, nonlinear geometry, choice of preconditioner, and other solution parameter.
Finally, the article ends with a summary of the main conclusions of the work.

@@

The governing equations are complemented by boundary conditions, with three types considered here: prescribed displacement, prescribed traction, and symmetry.
The definition of the engineering stress ($\bb{\sigma}_s$) and true stress ($\bb{\sigma}$) in Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL} is given by a chosen mechanical law.
Five mechanical laws are considered in this work, as briefly outlined in Appendix \ref{app:mechLaws}: linear elasticity (Hooke's law), St.\,Venant-Kirchhoff hyperelasticity, neo-Hookean hyperelasticity, Guccione hyperelasticity, and neo-Hookean $J_2$ hyperelastoplasticity.

@@

\paragraph{Case 6: Elastic plate behind a rigid cylinder}
The final test case, introduced by \citet{Turek2006}, extends the classic flow around a cylinder in a channel problem \citep{Ferziger2002} into a well-established fluid-solid interaction benchmark.
The \texttt{FSI3} variant of the case examined here (Figure \ref{fig:hronTurek-mesh}) consists of a horizontal channel (0.41 m in height, 2.5 m in length) with a rigid cylinder of radius 0.05 m, where the cylinder centre is 0.2 m from the bottom and inlet (left) boundaries.
A parabolic velocity is prescribed at the inlet velocity with a mean value of 2 m/s.
A St.\,Venant-Kirchhoff hyperelastic plate ($E = 5.6$ MPa, $\nu = 0.4$) of 0.35 m in length and 0.02 m in height is attached to the right-hand side of the rigid cylinder.
The fluid model is assumed to be isothermal, incompressible, and laminar and adopts the segregated PIMPLE solution algorithm.
The fluid's kinematic viscosity is 0.001 m$^2$/s and density is 1000 kg/m$^3$, while the solid's density is $10\,000$ kg/m$^3$.
The interface quasi-Newton coupling approach with inverse Jacobian from a least-squares model \cite{Degroote2009} is employed for the fluid-solid interaction coupling.
The fluid-solid interface residual is reduced by four orders of magnitude within each time step.
Further details of the fluid-solid interaction procedure are found in \citet{Tukovic2018a}.
%\begin{figure}[htbp]
%   \centering
%%	\subfigure[Geometry (taken from \citet{Tukovic2018a})]
%%	{
%	   \includegraphics[width=0.7\textwidth]{figures/hronTurek-geometry} 
%%   	}
%   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
%   \label{fig:hronTurek-geometry}
%\end{figure}
Three successively refined quadrilateral meshes are employed:
The fluid region meshes have $1\,252$, $5\,008$, and $20\,032$ cells, while the solid region meshes have 156,  624 and $2\,496$ cells.
The total simulation time is 20 s, and the time-step size is 0.5 ms. 
\begin{figure}[htbp]
   \centering
	\subfigure[Mesh containing $5\,336$ cells in the fluid region and 630 cells in the solid region, where a closeup of the plate is shown in the right]
	{
   		\includegraphics[width=0.7\textwidth]{figures/hronTurek-mesh}
	}
	\subfigure[Close-up of the mesh around the plate]
	{
   		\includegraphics[width=0.4\textwidth]{figures/hronTurek-mesh-zoom}  
   	}
   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
   \label{fig:hronTurek-mesh}
\end{figure}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Resource Requirements and Robustness}
\label{sec:resource_requirements}
%%--------------------------------------------------------------------------------------------------------------------%%
This section compares resource requirements and the robustness of the Jacobian-free Newton-Krylov and segregated approaches on the cases presented in Section \ref{sec:accuracy}.
Specifically, time and memory requirements are analysed, and whether the solution approach converged for all meshes and loading steps.
%In some cases, time and memory requirements from finite element software \hl{Abaqus} are given for reference.
In all cases, the Jacobian-free Newton-Krylov approach used the generalised minimal residual method (GMRES) linear solver with the direct LU preconditioner for 2-D and the Hypre \emph{BoomerAMG} multigrid preconditioner for 3-D, unless stated otherwise.
A restart value of 30 is used for the GMRES solver, where the \emph{loose} variant of the GMRES solver is used, which retains a small number of extra solution vectors across restarts (2 in this case).
In contrast, The segregated approach used the conjugate gradient linear solver and the Hypre \emph{BoomerAMG} multigrid preconditioner for all cases.
Time and memory requirements are implementation and hardware-specific; nonetheless, it is insightful to see the relative performances of the Jacobian-free Newton-Krylov and segregated approaches on the same hardware and within the same implementation framework.
Clock times and memory usage (measured with the GNU time utility) were generated using a Mac Studio with an M1 Ultra CPU on one CPU core, where the code was built with the Clang compiler (version 16.0.0).
%\hl{Present data for one hardware first, and give comparison after.}
Examination of multi-CPU-core parallelisation is left to Section \ref{sec:parallelisation}.
% where several mesh densities are examined.
%Time and memory requirements from the classic segregated approach are given for comparison, along with the requirements from commercial finite element software Abaqus (version XXX): \hl{ask Dylan to create the Abaqus results when we are happy the paper will not change}.
%The finite element approach uses a coupled solution algorithm and direct linear solver, where a Newton-Raphson method is used for resolving nonlinearities.
%\hl{Take care when comparing times between different hardware}

@@

Regarding memory usage, below approximately $50$ k degrees of freedom, the memory usage is the same for all three approaches (approximately $80$ MB), corresponding to the solver's minimum memory overhead.
For greater numbers of degrees of freedom, the LU approach uses the greatest amount of memory for both 3-D cases (idealised ventricle, elliptic plate), while the ILU(5) uses the least.
In contrast, the LU approach uses the smallest amount of memory for the 2-D case (membrane i) mesh densities examined.
The rate of memory increase is seen to be much steeper for the LU approach on the 3-D cases than the multigrid and ILU(5) approaches, with the LU results for the finest meshes not shown as they required more than the maximum available memory (64 GB).
%\begin{itemize}
%	\item \hl{Ventricle finesh mesh: ILU5 diverges and LU requires too much memory}
%	\item \hl{Elliptic plate finest mesh: LU requires too much memory}
%	\item \hl{Comment on the results: e.g. fastest, slowest, most/least iterations, etc}
%	\item \hl{Provide insights, if any: e.g. why is one faster or slower}
%	\item \hl{Conclusion: which preconditioner if 'best' for which situation: probably LU is best for 'small' cases and multigrid for 'large' cases}
%\end{itemize}
\begin{figure}[htbp]
   \centering
	\subfigure[Clock time]
	{
   		\includegraphics[width=0.48\textwidth]{figures/time}
	}
	\subfigure[Maximum memory]
	{
   		\includegraphics[width=0.48\textwidth]{figures/memory}  
   	}
   \caption{The clock times and maximum memory usage for three preconditioning approaches (multigrid, ILU(5), and LU) as a function of degrees of freedom}
   \label{fig:times_memory}
\end{figure}
%\begin{table}[htb]
%	\centering
%		\begin{tabular}{lllll}
%			\hline
%			Case & Number of Cells & Preconditioner & Time (in s) & Memory (in MB)  \\
%			\hline 
%			Case X & 1,234 & LU & 345 & 1,234  \\
%			Case X & 12,345 & LU & 345 & 1,234  \\
%			Case X & 123,456 & LU & 345 & 1,234  \\
%			Case X & 1,234 & ILU(0) & 345 & 1,234  \\
%			Case X & 12,345 & ILU(0) & 345 & 1,234  \\
%			Case X & 123,456 & ILU(0) & 345 & 1,234  \\
%			Case X & 1,234 & Multigrid & 345 & 1,234  \\
%			Case X & 12,345 & Multigrid & 345 & 1,234  \\
%			Case X & 123,456 & Multigrid & 345 & 1,234  \\
%			Case Y & 1,234 & LU & 345 & 1,234  \\
%			Case Z & 1,234 & LU & 345 & 1,234  \\
%			\hline
%		\end{tabular}
%	\caption{Execution times and maximum memory usage for different preconditioners on varying cases and mesh densities \hl{Try other value of N for ILU?}}
%	\label{tab:preconditioner}
%\end{table}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Parallelisation}
\label{sec:parallelisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this final analysis, the multi-CPU-core parallel scaling performance of the proposed Jacobian-free Newton Krylov method is compared with the segregated approach.
A \emph{strong} scaling study is performed, where the clock time to solve a fixed-size problem is measured as the number of CPU cores is increased.
The elliptic plate (3-D, linear elastic) mesh with $2\,438\,242$ cells is chosen for the scaling analysis.
Parallelisation adopts the domain decomposition approach, where the mesh is decompose into one sub-domain for each CPU core.
In the current work, the Scotch decomposition approach \citep{Pellegrini2012} is employed.
The Jacobian-free Newton Krylov approach uses the GMRES linear solver where the performance of the multigrid and LU preconditioners are compared.
The segregated approach uses the conjugate gradient solver with the incomplete (zero in fill) Cholesky preconditioner.
The performance of the LU preconditioner 
The cases are run on the Meluxina high-performance computing system, where each standard computing node contains $2\times$ AMD EPYC Rome 7H12 64c 2.6GHz CPUs with 512 GB of memory.
In the current study, the number of CPU cores are varied from 1 to $1\,048$ ($\left[ 1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1\,048 \right]$).
In the ideal case, the speedup should double when the number of cores is doubled; in reality, inter-CPU-core communication reduces the parallel scaling efficiency below the ideal.
